---
title: "p8105_hw5_fwt2107"
author: "Felix Tran"
date: "November 3, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "right"))
```

# Problem 1

1. Create function **tidy_subject_data()** which will be used later to tidy 
the subject data.  

2. Read in all file names

3. Read in the data contained in those files to form a new variable of lists in 
the tibble (each file's data is an individual dataset)

4. Remove the '.csv' from the file names to leave the subject number and 
treatment arm

5. Plot spaghetti plot of subject scores over time by treatment group
```{r}
tidy_subject_data <- function(df) {
    gather(data = df, key = 'week', value = 'score', week_1:week_8) %>% 
    separate(col = week, into = c('deleteable', 'week'), sep = '_') %>% 
    select(-deleteable)
}

study_df <- tibble(
  file_name = list.files('./hw5_data/data'),
  subject_data = purrr::map(str_c('./hw5_data/data/', file_name), 
                            readr::read_csv)
) %>% 
  separate(col = file_name, into = c('file_name', 'deletable'), sep = '\\.') %>% 
  select(-deletable) %>% 
  mutate(subject_data = purrr::map(subject_data, tidy_subject_data),
         treatment_group = 
           if_else(str_detect(file_name, 'con'), 'Control', 'Experimental'))
  
study_df %>% 
  unnest %>% 
  ggplot(aes(x = week, y = score, color = treatment_group, group = file_name)) +
  geom_line(alpha = 0.6) +
  facet_grid(. ~ treatment_group) +
  labs(title = 'Subject scores over time by treatment group',
       x = 'Week',
       y = 'Score')
```
At the start of the study, the experimental group scores ranged between -1 to 4, 
while the control group ranged between 0 to 2.5. The scores in the control group 
as a whole remained unchanged by the end of the study. However, the scores in 
the experimental group as a whole increased, placing the experimental group
between 2.5 to 7.5 points. 



# Problem 2

### Describing the raw data

The raw data contains 12 variables:

1. uid - A unique ID for each homicide including the city

2. reported_date - The date when the homicide was reported (YYMMDD)

3. victim_last - Victim's last name

4. victim_first - Victim's first name

5. victim_race - Victim's race

6. victim_age - Victim's age in years

7. victim_sex - Victim's sex 

8. City - City in which the homicide occurred

9. State - State in which the homicide occurred 

10. lat - Latitude where the homicide occurred

11. long - Longitude where the homicide occurred

12. disposition - Describes if the homicide investigation resulted in an arrest,
no arrest, or if the case was closed without an arrest being made
```{r}
data_url <- RCurl::getURL('https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv')

homicide_df <- readr::read_csv(data_url) %>% 
  mutate(victim_age = as.integer(victim_age),
         reported_date = lubridate::ymd(reported_date))
head(homicide_df)
unique(homicide_df$disposition)


prop_male <- round((length(which(homicide_df$victim_sex == 'Male')) / 
                nrow(homicide_df)) * 100, digits = 0)

prop_unsolved <- round((length(which(
  homicide_df$disposition != 'Closed by arrest')) / nrow(homicide_df)) * 100, 
  digits = 0)
```
The dataset contains information about `r nrow(homicide_df)` homicides which 
occurred in `r length(unique(homicide_df$city))` cities across 
`r length(unique(homicide_df$state))` states between 2007 - 2017. `r prop_male`%
of the victims were male, and the average victim age was 
`r round(mean(homicide_df$victim_age, na.rm = T), digits = 0)` years old. 
`r prop_unsolved`% of all homicides are unsolved (closed without arrest or 
open/no arrest).

### Homicides by city

1. Upon inspection of the data, one homicide listed Tulsa as the city but AL as
the state. This mistake was corrected by changing the reported state to OK.

1. Create variable **city_state** to concatenate city and state information

2. Count the total number of homicides and the number of unsolved homicides by
location
```{r}
homicide_df <- homicide_df %>% 
  mutate(state = if_else(state == 'AL' & city == 'Tulsa', 'OK', state)) %>% 
  mutate(city_state = str_c(city, ', ', state))

options(tibble.print_max = Inf)
homicide_df %>% 
  group_by(city_state) %>% 
  summarize(total_homicides = n(),
            unsolved_homicides = 
              length(which(disposition != 'Closed by arrest')))
```

### Baltimore, MD
1. Repeat the same process as the previous example, but focusing only on 
Baltimore

2. Use the **prop.test** function to estimate the proportion of homicides that
are unsolved in Baltimore with 95% CI.
```{r}
baltimore_results_df <- homicide_df %>% 
  group_by(city_state) %>% 
  filter(city_state == 'Baltimore, MD') %>% 
  summarize(unsolved_homicides = 
              length(which(disposition != 'Closed by arrest')),
            total_homicides = n(),
            prop_results = purrr::map2(unsolved_homicides, total_homicides,
                                       prop.test)) %>% 
  select(prop_results) %>% 
  broom::tidy(prop_results) %>% 
  janitor::clean_names()

balt_unsolved_est <- round(baltimore_results_df$estimate * 100, digits = 0)
balt_unsolved_conf_low <- round((baltimore_results_df$conf_low * 100), 
                               digits = 0)
balt_unsolved_conf_high <- round((baltimore_results_df$conf_high * 100), 
                                digits = 0)
```
An estimated `r balt_unsolved_est`% (95% CI: `r balt_unsolved_conf_low`%, 
`r balt_unsolved_conf_high`%) of homicides in Baltimore are unsolved. 


### Creating a dataset with the estimate and 95% CI of unsolved homicides in
### all 50 cities

1. Repeat the same process as the previous Baltimore example, but with all 
50 cities

2. Use the **prop.test** function to estimate the proportion of homicides that
are unsolved in each city with 95% CI.
```{r}
unsolved_df <- homicide_df %>% 
  group_by(city_state) %>% 
  summarize(unsolved_homicides = 
              length(which(disposition != 'Closed by arrest')),
            total_homicides = n(),
            prop_results = purrr::map2(unsolved_homicides, total_homicides,
                                       prop.test)) %>% 
  select(city_state, prop_results) %>% 
  mutate(prop_results = purrr::map(prop_results, broom::tidy)) %>% 
  unnest %>% 
  janitor::clean_names() %>% 
  select(city_state, estimate, conf_low, conf_high)
```


### Plotting the newly created dataset

1. Created a function **prop_transform** to transform the values of the 
estimated proportion and confidence limits for unsolved homicides into 
percentages

2. Saved the order of the 50 cities according to proportion of unsolved
homicides in variable **city_factor_order**

3. Transformed dataset by applying **prop_transform** to the estimated 
proportion and confidence limits, as well as transforming variable 
**city_state** into a factor with the values ordered according to 
**city_factor_order**

4. Plot the estimated proportion of unsolved homicides with 95% CI in the cities
```{r}
prop_transform <- function(x) {
  round((x * 100), digits = 0)
}

city_factor_order <- unsolved_df$city_state[order(unsolved_df$estimate)]

unsolved_df %>% 
  mutate(estimate = prop_transform(estimate),
         conf_low = prop_transform(conf_low),
         conf_high = prop_transform(conf_high),
         city_state = readr::parse_factor(city_state, city_factor_order)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), 
           fill = 'blue', alpha = 0.6) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), alpha = 0.75) +
  labs(title = 'Estimated proportion of homicides that are unsolved for 
       50 US cities with 95% confidence intervals',
       y = 'Estimate (%)',
       x = 'City') +
  coord_flip() +
  theme(axis.text.y = element_text(size = 3.5, color = 'black'))
```

The proportion of homicides which are unsolved range from less than 30% in 
Richmond, VA to over 70% in Chicago, IL. In the majority of the 50 listed 
cities, over half of all homicides are unsolved. 
